---
title: '`dplyr`: The Yotabites Demo'
author: "Shiv Sundar"
date: "June 5, 2018"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
worldCountries <- read.csv("worldCountries.csv")
library(dplyr)
knitr::opts_chunk$set(echo = TRUE)
```

#Introduction
This tutorial was created to inform the user about every function provided by the `dplyr` package. Commonly used and simple functions will include a short use case and demo. 

##The pipe operator
One of the most used tools provided by `dplyr` is the pipe (`%>%`) operator. We can use it to chain together multiple functions. It's kind of like saying: do this, **then** do this, **then** do this. We'll cover the use and syntax for it as we go.

##Commonly used `dplyr` functions

###Are these two data frames the same?
The `all_equal()` function is used to compare two data frames. One can use this function to check for equality without considering row order, column order, or both.

####Usage
`all_equal(df1, df2, ignore_col_order = TRUE, ignore_row_order = TRUE, convert = FALSE, ...)`

####Parameters
* `df1`, `df2`: dataframes to compare.
* `ignore_col_order`: should the ordering of columns be ignored?
* `ignore_row_order`: should the ordering of rows be ignored?
* `convert`: should the function attempt to coerce factors to characters and integers to doubles?
* `...`: **not used**. required to comply with `all.equal()` usage.

####Examples
```{r include=FALSE}
worldCountries1 <- sample_n(worldCountries[ , c("Country", "Population", "Phones..per.1000.")], 15)
worldCountriesMix <- worldCountries1[sample(nrow(worldCountries1)), sample(ncol(worldCountries1))]
```

Here we see what the data looks like.
```{r}
head(worldCountries1)
head(worldCountriesMix)
```

Here we test different variations of the function.
```{r}
all_equal(worldCountries1, worldCountriesMix)
all_equal(worldCountries1, worldCountriesMix, TRUE, FALSE)
all_equal(worldCountries1, worldCountriesMix, FALSE, FALSE)
```

###I want to sort my data by a specified order!
Data can be sorted by using the `arrange()` function. It sorts using by using the natural order of the data type. Factors are normally sorted by alphabetical order. To order them differently, please use the `levels()` function.

####Usage
`arrange(.data, ...)`

####Parameters
* `.data`: the dataframe to arrange
* `...`: the columns to sort by

####Examples
```{r}
head(arrange(worldCountries1, Population))
head(worldCountries1 %>%
  arrange(desc(Phones..per.1000.)))
```

We just saw the first use of the pipe operator! In English, we took the data frame `worldCountries1`, **then** arranged its data by the values in the `Phones..per.1000.` column. We can accomplish this task without using the pipe operator as well.

```{r}
head(arrange(worldCountries1, desc(Phones..per.1000.)))
```

Functions provided by `dplyr` support both options that we just covered. It's up to your preference. However, in this demo we will use the pipe operator to familiarize you with its use.

###Which values in a vector fall between 10.4 and 73?
How can we answer this question? Well, `dplyr` actually provides us with a function called `between()`. That's convenient! 

####Usage
`between(vector, min, max)`

####Parameters
* `vector`: the values to check
* `min`: the minimum value (inclusive)
* `max`: the maximum value (inclusive)

####Examples
```{r}
newData <- sample_n(worldCountriesMix, 15)$Population
head(newData)
newData%>%
  between(200000, 7500000)
```


##Datasets provided by `dplyr`
These datasets are bundled with the `dplyr` package for testing and manipulating.

* `band_members`
* `band_instruments`
* `band_instruments2`

##Other functions not covered

all_vars
arrange_all
as.table.tbl_cube
as.tbl_cube
auto_copy
between
bind
case_when
coalesce
compute
copy_to
cumall
desc
distinct
do
dr_dplyr
explain
filter
filter_all
funs
groups
group_by
group_by_all
ident
if_else
join
join.tbl_df
lead-lag
mutate
n
nasa
na_if
near
nth
n_distinct
order_by
pull
ranking
recode
rowwise
sample
scoped
select
select_all
select_vars
setops
slice
sql
src_dbi
starwars
storms
summarise
summarise_all
tally
tbl
tbl_cube
top_n
vars