---
title: 'Using tidyr: A Short Demo'
author: "Shiv Sundar"
date: "May 30, 2018"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
library(tidyr)
library(dplyr)
data <- read.csv("sales.csv")
knitr::opts_chunk$set(echo = TRUE)
```

#Introduction
In this demo, we will take a look at a package called `tidyr` and its most used functions. To do this, let's assume the role of 

##separate()
The data in the column titled `event_date_time` contains repeat data from the `event_date` column. We can remove this duplicate data by using the `separate()` command to split it into two columns.

```{r}
newData <- separate(data, event_date_time, c("event_dt", "event_time"), sep = " ")
head(newData$event_dt)
head(newData$event_time)
```

##unite()
Here, we put two columns together by using the `unite()` function. Our data originally looks like this.
```{r}
head(newData$venue_city)
head(newData$venue_state)
```

Now we combine the two columns 

```{r}
testData <- unite(newData, "venue_loc", venue_city, venue_state, sep = ", ")
head(testData$venue_loc)
```

##gather()
```{r}
testData <- spread(testData, timezn_nm, venue_loc)
testData <- select(testData, seq(ncol(testData) - 3, ncol(testData)))
head(testData)
```

##spread()

