---
title: 'Building web applications with `shiny` and RStudio Connect'
author: "Shiv Sundar"
date: "June 18, 2018"
runtime: shiny
output: html_document

---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(shiny)
```

#Introduction
In this demo, we will cover how to use `shiny` and the differences between RStudio Connect and the [shinyapps.io](https://www.shinyapps.io) service.

##`shiny`
`shiny` is a package used to quickly build web applications.

###Accessing data in `shiny`
`shiny` stores all objects in a vector called `input`. For example, if one creates an input object with the object id `button2`, then that object's value can be accessed by calling `input$button2`.

###Types of inputs available in Shiny
In the `shiny` package, there are 14 total UI input components that one can use. These are:

* `actionButton()`
* `actionLink()`
* `checkboxGroupInput()`
* `checkboxInput()`
* `dateInput()`
* `dateRangeInput()`
* `fileInput()`
* `numericInput()`
* `passwordInput()`
* `radioInput()`
* `selectInput()`
* `sliderInput()`
* `submitButton()`
* `textInput()`

`shiny` also has many output components. These can be 

For example, here we can see the implementation of a application that displays the selected date. Selecting a date updates the text to display what was selected.

```{r, eval = FALSE}
dateInput("calendar", label = "Choose date to display")
renderText({paste("Selected Date: ", as.character(input$calendar), sep = "")})
```

```{r, echo = FALSE}
dateInput("calendar", label = "Choose date to display")
renderText({paste("Selected Date: ", as.character(input$calendar), sep = "")})
```

RStudio Connect is a service that does more than `shinyapps.io`. While `shinyapps.io` is used to quickly get an app on the web, RStudio Connect is aimed at building a full-fledged website and 

##What is `shinyapps.io`?
`shinyapps.io` is a service that aims to get your interactive web application published as soon as possible. It only offers support for applications and has limited authentication features. 

##What is RStudio Connect?
RStudio Connect is a service provided by RStudio. It is focused on publishing multiple file types, including Shiny applications and R Markdown reports enhanced with Shiny.

##What are the differences between the two?
`shinyapps.io` | RStudio Connect
-------------- | --------------
Support for Shiny apps only | Support for Shiny apps, RMarkdowns, and more
Focused on simple development | Provides a robust framework for publishing
Used by smaller applications | Used by academic institutions and companies
Free tier exists | Basic service costs $15,000
