---
title: 'Building web applications with `shiny` and RStudio Connect'
author: "Shiv Sundar"
date: "June 18, 2018"
runtime: shiny
output: html_document

---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(shiny)
```

#Introduction
In this demo, we will cover how to use `shiny` and RStudio Connect, and the purpose of each.

##`shiny`
###Types of inputs available in Shiny
In the `shiny` package, there are 14 total UI components that one can use. These are:

* `actionButton()`
* `actionLink()`
* `checkboxGroupInput()`
* `checkboxInput()`
* `dateInput()`
* `dateRangeInput()`
* `fileInput()`
* `numericInput()`
* `passwordInput()`
* `radioInput()`
* `selectInput()`
* `sliderInput()`
* `submitButton()`
* `textInput()`

For example, here we can see the implementation of a application that displays the selected date.
```{r, eval = FALSE}
dateInput("calendar", label = "Choose date to display")
renderText({as.character(input$calendar)})
```

```{r, echo = FALSE}
dateInput("calendar", label = "Choose date to display")
renderText({as.character(input$calendar)})
```

##RStudio Connect
###What is RStudio Connect?
RStudio Connect is a service that does more than `shiny`.